A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-23 doc/paper.tex
Redesign of limitations.

	\begin{itemize} \item API improvement \item Detection of code smells \item
		Spreadsheet education \end{itemize}

@@

	\subsection{API Improvement}

@@

	\subsection{Study Limitations}

	Still, several limitations cropped up during the study that complicate the
	evaluation. For one thing, though the source of the data was not explicitly
	stated at the outset of the session, several users studied the file names
	(unchanged as they were) to deduce its roots in finance, if not Enron itself.
	Because of these notorious connotations, then, some were inclined from this
	moment of discovery to orient their exploration around finance-related
	functions, potentially limiting their findings or perceived users for the tool
	-- though some also said that, without this essential context, they could not
	have properly explored a dataset in the first place. Comments like these point
	out the trade-offs of unguided exploration, too: though it might not restrict
	them to a single purpose, it might, at the same time, not afford them any
	mindset at all with which to interpret the data. Another potential problem is
	that the data which users explored was processed and created before we had
	fully completed the tool, meaning some of the counts and patterns in the trees
	were inaccurate. However, we decided that this was not a pressing concern,
	being that we were primarily with whether any interesting reports could be made
	at all, not with whether these reports were 100\% accurate assessments of
	reality. \par

	Throughout the study, it became apparent that some of the concepts,
	particularly optional arguments, were easily lost in the representation. To
	illustrate this, consider the following two trees.	
	The difference is, perhaps, too subtle: on the left, the tree contains
	instances of AND functions with any number of arguments, up to four; on the
	right, the tree contains only instances with exactly four arguments. In other
	words, an instance of "=AND(A1, B2, C3, D4)" would be found in both trees,
	while "=AND(A1, B2)" would be found in only the left. Though the users
	naturally did not explicitly report misinterpreting the icons, it became clear
	in their out-loud thoughts that they viewed each box as a fundamentally
	different set of possibilities in formula construction rather than just a
	specific parameter. Though we clarified the meanings when these
	misunderstandings became apparent, these events point to a failure in the tool
	to properly indicate every function with clarity.   \par

	\section{Limitations} Beyond the negative responses in the user study, a few
	more inherent limitations beset the tool. These largely resulted from
	trade-offs in which we found no acceptable way to win the best of both worlds,
	and necessary compromises were made. \par

	Because the data collection depends on POI's formula parser, it affords no
	leeway or partial information from a formula; it either processes it perfectly
	or throws it out. As such, the visualization inhibits insight into anything
	with syntactical errors or third-party functions that can't be evaluated
	without additional tools. Note, however, that this does not include standard
	Excel errors like \#REF! and \#DIV/0, which the parser handles well. This poses
	more egregious problems, however, when POI does not support the parsing of
	certain functions, such as EOMONTH (as obscure as it may be), preventing the
	healthy growth of those trees. \par

	Many functions have an expected order of arguments, but some, like SUM and
	MATCH, can be reordered in various ways and maintain their value. However, the
	representation that this tool uses reinforces for all functions that the order
	is important, even in these exceptions. Future work could design a new family
	of tree to accommodate these loose cases, however, by bucketing
	order-insignificant arguments together; this, of course, would still have to
	address the profusion of co-presented data. \par

	A larger and related problem is the loss of argument combination. For example,
	figure \ref{fig:index} shows the situation of the INDEX tree, wherein each
	argument position has a few distinct possibilities. While it is easy to break
	down, for each position, which argument types are most common (ranges, MATCH,
	and numbers, respectively), the tree is silent on whether MATCH functions are
	used together in the second and third arguments, among other concerns. Though
	these argument combinations could be pieced together by manually inspecting the
	lists of examples, it is nevertheless inconvenient, though not outside the
	realm of future improvement.

	A threat to the exploratory philosophy may also mount from the problem of too
	much data, particularly in the popular functions, like SUM and IF. We had added
	some precautions before the study: for any position in the tree, for example,
	only the ten most common functions were displayed without clicking on the
	expansion arrow \includegraphics[scale=.35]{arrow}. Nevertheless, P3 remarked
	that they specifically avoided the most popular functions, anticipating a flood
	of nodes from which they could salvage nothing interesting -- which might not
	have been a bad prediction, considering that the SUM is built from 910 nodes
	and IF, at least 18000 [double-check these with refined data]. P2 corroborated
	this by explaining how they gravitated toward specific examples over the plane
	of nodes but complicates it further by also deeming the tool at its best when
	it shows either every possibility of an side-by-side (for comparison) or
	nothing on that branch at all, a balance that is hard to strike with node-dense
	trees. \par

	%To be sure, though, there is still plenty of work to improve this tool. One
	% activity
	%it does not support well now is that of comparison: for all of it's ability to
	% spit
	%out numbers and quantify patterns,

@@

>>>>

	\begin{itemize} \item Detection of code smells \item
		Spreadsheet education \end{itemize}

@@

@@

	\subsection{Study Limitations}
	Though we did not start the study by telling that the data source came from Enron, 
	two users studied the file names
	to deduce its roots in finance, if not Enron itself.
	Because of these notorious connotations, then, they might have been inclined from this
	information to orient their exploration around finance-related
	functions, potentially limiting their findings or perceived users for the tool. Though,
	one also said that, without this essential context, they could not
	have properly explored a dataset in the first place. Comments like these point
	out the trade-offs of unguided exploration, too: though it might not restrict
	them to a single purpose, it might, at the same time, not afford them any
	mindset at all with which to interpret the data. \par

	Furthermore, the data which users explored was processed and created before we had
	fully completed the tool, meaning some of the counts and patterns in the trees
	were inaccurate. However, we decided that this was not a pressing concern,
	being that we were primarily concerned with whether any interesting reports could be made
	at all, not with whether these reports were 100\% accurate assessments of
	reality. \par

	Throughout the study, we noticed that users misunderstood some of the icons
	and, as such, their early discoveries were inaccurate. For example, P1 interpreted
	the numbers on the argument boxes to mean that all of the box's children had exactly
	that many arguments -- that is, clicking on the dark blue ``1" would return all the single-argument functions
	instead of the first argument of a function -- and P2 initially took the solid lines to mean
	different paths in formula construction rather than different arguments of the same function. 
	Though the users naturally did not explicitly report misinterpreting
	the icons, these misunderstandings became clear in their early evaluations of the
	spreadsheet practice. Though we clarified the meanings when
	these misunderstandings became apparent, these events point to a failure in the
	tool to properly indicate every function with clarity.   \par

	\section{Discussion} 	
	We began these studies with the intention of evaluating our progress towards the
	goals in section \ref{goals}. Working from these results, we noted many places
	where the tool aligned well or fell short.

	\subsection{Unbiased Exploration}	
	The first goal -- that of full exploration -- is threatened by the presence of certain
	formulae that the tool does not display, such as those with nonstandard functions. 
	Because the data collection depends on POI's formula parser, it does not
	allow for any partial parsing of a formula; it either processes a formula perfectly
	or throws it out. As such, the visualization does not display anything
	with syntactical errors or third-party functions that POI cannot parse. This poses
	larger problems, however, when POI does not support the parsing of
	certain functions, such as EOMONTH, preventing the
	healthy growth of those trees. \par

	Participants also asked for more options to quickly navigate the tree. For example,
	P2 asked for the ability to expand the entire tree at once and icons that better
	communicated the presence of optional arguments and their quantities. The more 
	familiar participants grew with the tool, the faster they wanted to explore, yet
	it supported expert navigation up to the point of expanding all of a node's children
	up to one level. \par

	Other than that, we did avoid designing anything that intentionally evaluated
	formula designs. Nevertheless, \toolname inadvertently emphasizes bad design of
	certain types, such as all of those in section \ref{badsmells} marked by their
	visual signatures. Even so, our participants found the discovery of bad smell
	to be the most interesting aspect about it, to the point where they could
	imagine business scenarios in which micromanagers use it to ensure
	precise design parameters across a department's spreadsheets! To design in that
	direction, however, would mean a reevaluation of this design goal, since
	intentionally guiding users toward known smells may guide them away from
	new, unmarked ones.

	\subsection{Uncluttered Quantification}

	We realized early that some of the popular functions would present a problem: 
	some would have thousands of variations in large datasets, and showing all of
	these variations at once would hopelessly overwhelm the user. We therefore
	made a number of choices, such as the expansion arrows and the use of generic
	types in place of specific primitive values, that would reduce the variation
	to something more manageable.

	Nevertheless, during the user study, P3 remarked that they specifically avoided
	the most popular functions, anticipating a flood of nodes from which they could
	salvage nothing interesting -- which might not have been a bad prediction,
	considering that the SUM is built from 910 function and argument nodes stemming
	the root and IF, at least 18000. P2 corroborated this by explaining how they
	gravitated toward exploring specific examples in particular value nodes over
	navigating the full crowd of the tree, but they complicate it further by also
	saying that the tool at its best when it shows either every possibility of an
	argument side-by-side (for comparison) or nothing on that branch at all, a
	balance that is hard to strike with node-dense trees.

	A threat to our goals, then, arises from the problem of too much data,
	particularly in the popular functions, like SUM and IF. Despite our
	precautions, balancing the desire to show every single possibility of design
	with the desire not to overwhelm remains a challenge. \par

	\subsection{Informative Qualification}

@@

	A larger and related problem is the loss of argument combination. For example,
	figure \ref{fig:index} shows the situation of the INDEX tree, wherein each
	argument position has a few distinct possibilities. While it is easy to break
	down, for each position, which argument types are most common (ranges, MATCH,
	and numbers, respectively), the tree is silent on whether MATCH functions are
	used together in the second and third arguments, among other concerns. Though
	these argument combinations could be pieced together by manually inspecting the
	lists of examples, it is nevertheless inconvenient, though not outside the
	realm of future improvement. \par

	Further contextual information.

	Redesign.
